---
title: "Tigris package and census data"
author: "Mark Hagemann"
date: "September 21, 2017"
output: html_document
---

It's past 11 on a week night and I just got an idea of something I'd like to make. I want to look at census data and housing prices over time and space. I'll start with my new hometown of Richmond, VA.

I'll be using census TIGER data via the `tigris` R package.

```{r}
install.packages("tigris")
library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
```

I'll try looking at "census tract" data. This is not the most granular available, but I'm told it is a rough approximation of neighborhoods. 

```{r}
tigris::counties()
tracts <- tigris::tracts(state = "VA", county = "Richmond city")
tracts <- as(tracts, "sf")
# tracts2 <- tracts(state = "VA", county = "Richmond County")

class(tracts)
plot(tracts)
glimpse(tracts)
dim(tracts)
```


## Housing data

I also got housing price data from Zillow, at the zip-code scale. Let's see what that looks like. 

```{r}
zillow <- read.csv("data/Neighborhood_MedianValuePerSqft_AllHomes.csv")
names(zillow)
zillow[1:10, 1:9]
```

Neat! 

```{r}
zillow %>% 
  filter(Metro == "Richmond", State == "VA") %>% 
  glimpse()
```

Zillow shapefiles are 

```{r}
list.files("data")

zilhoods <- st_read("data/ZillowNeighborhoods-VA.shp") %>% 
  filter(City == "Richmond")

glimpse(zilhoods)
glimpse(tracts)

plot(zilhoods["Name"])
plot(tracts["NAME"])
```

I'm going to have to get more granular. Drilling down to the block level. This will omit 1990 as a data point, and will start instead at 2000.

```{r}
blocks <- tigris::blocks("VA", "Richmond city")
blocks <- as(blocks, "sf")
st_write(blocks, dsn = "data/rva_blocks.shp")
```

